---
title: "Accessing Network Data from a Database"
author: "Christopher Callaghan - Common Operational Research Environment (CORE) Lab"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
options(width = 150)
```

Load the relevant libraries:

```{r, message=FALSE}
library(RSQLite)
library(dplyr)
library(dbplyr)
```

# Getting Up and Running

Set up access variables for the database:

```{r}
con <- dbConnect(SQLite(), "my-db.sqlite")
```


# Interfacing with the Portable Database

Retrieve the names of available tables:

```{r}
dbListTables(con)
```

You can use R to select variables from the tables in the database, like so:

```{r}
con %>%
  tbl("criminal_events") %>%
  filter(FBI.Code == "08B") %>%
  collect() %>%
  glimpse()
```

```{r}
con %>%
  tbl("person_info") %>%
  filter(Sex == "M") %>%
  collect() %>%
  glimpse()
```

```{r}
con %>%
  tbl("person_x_events") %>%
  collect() %>%
  glimpse()
```

However, you can also run these queries in SQL, using the code below:
```{r, message = TRUE}
con %>%
  tbl("criminal_events") %>%
  filter(FBI.Code == "08B") %>%
  show_query()
```

```{r, message = TRUE}
con %>%
  tbl("person_info") %>%
  filter(Sex == "M") %>%
  show_query()
```

```{r, message = TRUE}
con %>%
  tbl("person_x_events") %>%
  show_query()
```

You can get as creative as you want.


# Producing an Edge List

This section presents you with various ways to generate an edge list:

## Based on Crime Type

Let's get a person-to-event edge list for ORA, for `BATTERY` and `ASSAULT`:

```{r}
con %>%
  tbl("criminal_events") %>%
  filter(Primary.Type == "Battery" | Primary.Type == "Assault") %>%
  select(eventId, Primary.Type) %>%
  collect() -> wanted_events

con %>%
  tbl("person_x_events") %>%
  filter(eventId %in% !!wanted_events[['eventId']]) %>%
  select(eventId, PID, Primary.Type) %>%
  collect() %>%
  mutate(source_type = "Event",
         target_type = "Agent") %>%
  rename(source = eventId, target = PID, relationship = Primary.Type) %>%
  select(source_type, source, target_type, target, relationship) -> edge_list
```
```{r, echo = FALSE}
edge_list %>%
  glimpse()
```

Which you could translate to SQL using the `show_query()`.

## Based on Person Identity

Another way to query would be if you had a set of personal identifiers (`PID`) that you are interested in:

```{r}
wanted_ids <- c(2, 87, 225)

con %>%
  tbl("person_x_events") %>%
  filter(PID %in% wanted_ids) %>%
  select(eventId, PID, Primary.Type) %>%
  collect() %>%
  mutate(source_type = "Event",
         target_type = "Agent") %>%
  rename(source = eventId, target = PID, relationship = Primary.Type) %>%
  select(source_type, source, target_type, target, relationship) -> edge_list
```
```{r, echo = FALSE}
edge_list %>%
  glimpse()
```

## Based on Geography

Alternatively, you can isolate events on interest from an area:

```{r}
wanted_districts <- c(6, 7, 8)

con %>%
  tbl("criminal_events") %>%
  filter(District %in% wanted_districts) %>%
  select(eventId, Primary.Type) %>%
  collect() -> wanted_events

con %>%
  tbl("person_x_events") %>%
  filter(eventId %in% !!wanted_events[['eventId']]) %>%
  select(eventId, PID, Primary.Type) %>%
  collect() %>%
  mutate(source_type = "Event",
         target_type = "Agent") %>%
  rename(source = eventId, target = PID, relationship = Primary.Type) %>%
  select(source_type, source, target_type, target, relationship) -> edge_list
```
```{r, echo = FALSE}
edge_list %>%
  glimpse()
```

# Generate Node List

For any of the edge lists above, you way want to generate a node table. First, for the people:

```{r}
con %>%
  tbl("person_info") %>%
  filter(PID %in% !!edge_list[['target']]) %>%
  collect() -> nodes_agents
```
```{r, echo = FALSE}
nodes_agents %>%
  glimpse()
```

Second, for the organizations:

```{r}
con %>%
  tbl("criminal_events") %>%
  filter(eventId %in% !!edge_list[['source']]) %>%
  collect() -> nodes_organizations
```

```{r, echo = FALSE}
nodes_organizations %>%
  glimpse()
```
